\bf A Hierarchical Model for Heterogenous Reliability Field Data

### 2018-05-19 paper/m-lb-m-technometrics.tex
Fixes small equation typos

Following the suggestions given in \cite{LiMeeker2014} and Section 15.2 of \cite{intervals} we use an alternative parameterization
where the usual scale parameter $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
[-\log(1-p)\right ]^{\sigma}$ (which is also a scale parameter). Replacing $\alpha$ in (1) with the
$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
\begin{align*}
\nonumber
\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
\right ]\\
\label{equation:reparameterized.weibull.cdf}
&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0.
\end{align*}
There are three important reasons for using this parameterization.
\begin{itemize}
\item
Especially with a high-reliability product, it will be easier to
elicit prior information about a quantile  ($t_p$) in the lower tail of the
distribution than it will be to elicit prior information about
$\alpha$ (approximately the $0.63$ quantile). In addition, there is generally available
information about the shape parameter, $\sigma$, for a given failure
mechanism (e.g., if the failure is due to a wearout mechanism, then
it is known that $\sigma<1$). 
\item
Because of heavy censoring in reliability field data, the parameter
estimates of the $\mu$ and $\sigma$ parameters will generally be
highly correlated and thus specification of independent marginal
prior distributions would be be inappropriate. On the other hand,
estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
value of $p$, will be approximately independent, allowing the easier
elicitation and specification of independent marginal prior
distributions. For example, if a typical field data set has 10\% of units
failing, then choosing $t_{0.05}$ would 
work well.
\item Bayesian MCMC estimation will be better behaved due to reduced correlation between $t_p$ and $\sigma$ (relative to $\alpha$ and $\sigma$).
\end{itemize}

\subsection{Generalized Limited Failure Population model}
\label{subsec:GLFP model}
Let $F_1,F_2$ be Weibull distributions with parameters $(t_{p1},\sigma_1)$ and $(t_{p2}, \sigma_2)$, respectively.
The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, t_{p1},\sigma_1,t_{p2},\sigma_2)$. Then
$$\Pr(T \le t) = H(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$

The GLFP model can be understood as a mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$. $\delta_i$ is an indicator for whether unit $i$ is defective or not (i.e., susceptible to an early failure).  Here $F_{1}(t)$ is the cdf for the early failures and $F_{2}(t)$ is
the cdf for the wearout failures. Expressed conditional on~$\delta_i$,
\begin{align*}
P(T\le t | \delta_i=1) =& 1 -(1-F_1(t))(1-F_2(t))\\
P(T\le t | \delta_i=0) =& F_2(t).
\end{align*}

@@

Taking the derivative of the (marginal) cdf, the density for the GLFP model is
\begin{align*}
h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_1(t|t_{p1},\sigma_1\right).
\end{align*}

@@

When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the traditional parameters of the distribution.  Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions (See the Supplementary Material for standard density plots of the priors).  Using this convention the prior distributions used in our analyses are
%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,\,7.6\times 10^6 \rangle,\quad
t_{0.2,2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
$$\pi \sim \op{Logit-normal}\langle 0.001,\, 0.71 \rangle,\quad
\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 0.0074 ,\,130 \rangle. $$
%\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$

@@

In a hierarchical model, the parameters are modeled as random variables, varying across different sub-populations.  For example
\begin{equation*}
\sigma_{1g} \ind \op{Lognormal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
\end{equation*}
\begin{equation*}
\sigma_{2g} \ind \op{Lognormal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{Tr}\left(0, 1\right)
\end{equation*}
\begin{equation}
\label{eq:hier-model}
t_{p_{1}g} = \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Lognormal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
\end{equation}
\begin{equation*}
t_{p_{2}g} = \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Lognormal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
\end{equation*}
\begin{equation*}
\pi_g \ind \op{Logit-normal}(\eta_\pi, \tau^2_\pi).
\end{equation*}

@@

\begin{align*}
\mbox{lpd} &= \sum_{g=1}^G \sum_{i=1}^{n_g} \log[p(t_{new,g,i}|t, t^L, c)] \\
&= \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right].
\end{align*}
\noindent Using this criterion, we should choose the model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$. Notationally suppressing the conditioning on $t^L$ and $c$,
\begin{equation}
\mbox{elpd} = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i},
\label{elpd}
\end{equation}
where $h$ is a density for the true distribution of $t_{new,g,i}$. Although $h$ is unknown, leave-one-out cross-validation provides a robust estimator:
\begin{equation}
\widehat{\text{elpd}} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})].
\label{loo-est}
\end{equation}
The \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent. It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $p(t_{g,i}|t_{-(g,i)})$ with $\sum_{s=1}^S w_{i,s} p(t_{g,i})/\sum_{s=1}^S w_{i,s}$. In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result.  A measure of the effective number of parameters is also available using the sum of the posterior variance of the log predictive density for each observation.
%When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically. 
% To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.  As outlined in Vehtari et. al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters rather than re-fitting the model with different data sets \cite{vehtari}.  For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with data point $i$ removed.  
% The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).

@@

\paragraph{Model 2} $t_{p_{2}g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the early failure mode:
$$ \pi \sim \op{Logit-normal}\langle 0.007,\, 0.26 \rangle,$$
$$\sigma_1 \sim \op{Lognormal}\langle 0.14,\, 7.1\rangle,$$ 
$$t_{0.5} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle.$$

@@

\begin{enumerate}
\item Sample $r$ uniformly from $\{1,\ldots,S\}$.\\ For $i=1,\ldots,n_g$, do:
\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(r)},t_{p1}^{(r)},\sigma_1^{(r)},t_{p_{2}g}^{(r)}, \sigma_{2g}^{(r)})$.
\item Repeat step 2 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$.
\item Set $d_{g,i}^{(s)}= \begin{cases} \max \{t_{g,i}: i \ge 1\}, & \mbox{ if }c_{g,i}=0\\
t_{g,i}, & \mbox{ if }c_{g, i}=1. \end{cases}$\\
Lastly, set $(t_{rep,g,i}^{(s)},c_{rep,g,i}^{(s)}) = \begin{cases}
  (d_{g,i}^{(s)},1), & \mbox{ if } \tilde{t}_{rep,g,i}^{(s)}>d_{g,i}^{(s)}\\
  (\tilde{t}_{rep,g,i}^{(s)},0), & \mbox{ if }\tilde{t}_{rep,g,i}^{(s)} \le d_{g,i}^{(s)}.
\end{cases}$
\end{enumerate}
Finally, plot K-M estimates based on the $R$ replicated data sets along with the one based on the observed data on the same axes. Discrepancies between the replicated and actual data can be used to diagnose inadequacies of the model.

@@

The mean time-to-failure (MTTF) for drive-model $d$ can be estimated by
$$\frac{1}{S} \sum_{s=1}^{S} t_{g,new}^{(s)}.$$
where $t_{g,new}^{(s)}$ is a draw from the posterior predictive distribution for drive-model $g$.

@@

Let $L(t) = e^{-0.4t}$ represent the value at the time of failure
relative to a new unit. We can rank the drive-models by 
$$\op{E}(L(t_{g,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{g,new}^{(s)}).$$

>>>>

Following the suggestions given in \cite{LiMeeker2014} and Section 15.2 of \cite{intervals} we use an alternative parameterization
where the usual scale parameter $\alpha$ is replaced by the $p$ quantile $t_{p}=\alpha \left
[-\log(1-p)\right ]^{\sigma}$ (which is also a scale parameter). Replacing $\alpha$ in (1) with the
$\alpha=t_{p}/[-\log(1-p)]^{\sigma}$ gives
\begin{align*}
\nonumber
\Pr(T \leq t|t_{p},\sigma ) = F(t|t_{p},\sigma)&=1-
\exp \left [-\left (\frac{t}{t_{p}/[-\log(1-p)]^{\sigma}} \right )^{1/ \sigma}
\right ]\\
\label{equation:reparameterized.weibull.cdf}
&=1 - \exp\left[\log(1-p)\left(\frac{t}{t_{p}}\right)^{1/\sigma}\right], \quad t > 0
\end{align*}
There are three important reasons for using this parameterization.
\begin{itemize}
\item
Especially with a high-reliability product, it will be easier to
elicit prior information about a quantile  ($t_p$) in the lower tail of the
distribution than it will be to elicit prior information about
$\alpha$ (approximately the $0.63$ quantile). In addition, there is generally available
information about the shape parameter, $\sigma$, for a given failure
mechanism (e.g., if the failure is due to a wearout mechanism, then
it is known that $\sigma<1$). 
\item
Because of heavy censoring in reliability field data, the parameter
estimates of the $\mu$ and $\sigma$ parameters will generally be
highly correlated and thus specification of independent marginal
prior distributions would be be inappropriate. On the other hand,
estimates of $t_{p}$ and $\sigma$, for some appropriately chosen
value of $p$, will be approximately independent, allowing the easier
elicitation and specification of independent marginal prior
distributions. For example, if a typical field data set has 10\% of units
failing, then choosing $t_{0.05}$ would 
work well.
\item Bayesian MCMC estimation will be better behaved due to reduced correlation between $t_p$ and $\sigma$ (relative to $\alpha$ and $\sigma$).
\end{itemize}

\subsection{Generalized Limited Failure Population model}
\label{subsec:GLFP model}
Let $F_1,F_2$ be Weibull distributions with parameters $(t_{p1},\sigma_1)$ and $(t_{p2}, \sigma_2)$, respectively.
The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, t_{p1},\sigma_1,t_{p2},\sigma_2)$ Then
$$\Pr(T \le t) = H(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$

The GLFP model can be understood as a mixture model with a binary latent variable, $\delta_i\ind \op{Bernoulli}(\pi)$. $\delta_i$ is an indicator for whether unit $i$ is defective or not (i.e., susceptible to an early failure).  Here $F_{1}(t)$ is the cdf for the early failures and $F_{2}(t)$ is
the cdf for the wearout failures. Expressed conditional on~$\delta_i$,
\begin{align*}
P(T\le t | \delta_i=1) &=  1 -(1-F_1(t))(1-F_2(t))\\
P(T\le t | \delta_i=0) &=  F_2(t)
\end{align*}

@@

Taking the derivative of the (marginal) cdf, the density for the GLFP model is
\begin{align*}
h(t|\pi, t_{p1},\sigma_1, t_{p2}, \sigma_2) = \pi & f_1(t|t_{p1},\sigma_1)\left(1-F_2(t|t_{p2},\sigma_2)\right) + \\ & f_2(t|t_{p2},\sigma_2)\left(1-\pi F_1(t|t_{p1},\sigma_1\right)
\end{align*}

@@

When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the traditional parameters of the distribution.  Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions (See the Supplementary Material for standard density plots of the priors).  Using this convention the prior distributions used in our analyses are
%$$t_{0.5,1} \sim \op{Lognormal}(8,4), \quad t_{0.2,2} \sim \op{Lognormal}(10,4)$$
$$t_{0.5,1} \sim \op{Lognormal} \langle 1.7,\,7.6\times 10^6 \rangle,\quad
t_{0.2,2} \sim \op{Lognormal}\langle 8.6, \, 5.6\times 10^7 \rangle$$
$$\pi \sim \op{Logit-normal}\langle 0.001,\, 0.71 \rangle,\quad
\sigma_1, \sigma_2 \stackrel{ind.}{\sim} \op{Lognormal}\langle 0.0074 ,\,130 \rangle $$
%\quad \sigma_2 \sim \op{Lognormal}\langle 7.4 \times 10 ^{-3} ,1.3 \times 10^2  \rangle $$

@@

In a hierarchical model, the parameters are modeled as random variables, varying across different sub-populations.  For example
\begin{equation*}
\sigma_{1g} \ind \op{Lognormal} \left( \eta_{\sigma_1}, \tau^2_{\sigma_1} \right)
\end{equation*}
\begin{equation*}
\sigma_{2g} \ind \op{Lognormal} \left( \eta_{\sigma_2}, \tau^2_{\sigma_2}\right)\op{Tr}\left(0, 1\right)
\end{equation*}
\begin{equation}
\label{eq:hier-model}
t_{p_{1}g} = \exp\left(\mu_{1g} + \sigma_{1g}\,\Phi^{-1}(p_1)\right)  \ind \op{Lognormal} \left(\eta_{t_{p_1}}, \tau^2_{t_{p_1}}\right)
\end{equation}
\begin{equation*}
t_{p_{2}g} = \exp\left(\mu_{2g} + \sigma_{2g}\,\Phi^{-1}(p_2)\right)  \ind \op{Lognormal} \left(\eta_{t_{p_2}}, \tau^2_{t_{p_2}}\right)
\end{equation*}
\begin{equation*}
\pi_g \ind \op{Logit-normal}(\eta_\pi, \tau^2_\pi)
\end{equation*}

@@

\begin{align*}
\mbox{lpd} &= \sum_{g=1}^G \sum_{i=1}^{n_g} \log[p(t_{new,g,i}|t, t^L, c)] \\
&= \sum_{g=1}^G \sum_{i=1}^{n_g} \log \left[ \int p(t_{new,g, i}|\theta) p(\theta|t,t^L,c) d\theta \right]
\end{align*}
\noindent Using this criterion, we should choose the model with the highest expected lpd (elpd) for a new data set $\{t_{new,g,i}:g=1,\ldots,G; i=1,\ldots,n_g\}$. Notationally suppressing the conditioning on $t^L$ and $c$,
\begin{equation}
\mbox{elpd} = \op{E}_h \mbox{lpd} = \sum_{g=1}^G \sum_{i=1}^{n_g} \int \log [p(t_{new, g, i}|t)] h(t_{new,g,i}) d t_{new,g,i}
\label{elpd}
\end{equation}
where $h$ is a density for the true distribution of $t_{new,g,i}$. Although $h$ is unknown, leave-one-out cross-validation provides a robust estimator:
\begin{equation}
\widehat{\text{elpd}} = \sum_{g=1}^G \sum_{i=1}^{n_g} \log [p(t_{g,i}|t_{-(g,i)})]
\label{loo-est}
\end{equation}
The \texttt{R} package, \texttt{loo} \citep{loo}, computes an approximation of (\ref{loo-est}) provided that the observations $t_i$ are conditionally independent. It employs an importance sampling method that uses $S$ saved MCMC draws to compute smoothed importance weights, $w_{i,s}$, approximating $p(t_{g,i}|t_{-(g,i)})$ with $\sum_{s=1}^S w_{i,s} p(t_{g,i})/\sum_{s=1}^S w_{i,s}$. In addition to providing an estimate of (\ref{elpd}), their method also produces a standard error based on an asymptotic result.  A measure of the effective number of parameters is also available using the sum of the posterior variance of the log predictive density for each observation.
%When $n$ is large, the distribution of the elpd is approximately normal and different models can be compared statistically. 
% To statistically compare the GLFP Models, we estimated the predictive accuracy of the 4 different models using an approximate leave-one-out cross validation (LOO) method.  As outlined in Vehtari et. al, the predictive accuracy from a fitted Bayesian model can be estimated simply with posterior draws of the model parameters rather than re-fitting the model with different data sets \cite{vehtari}.  For each observation, $i$, the log point-wise predictive density is calculated over the full set of posterior samples with data point $i$ removed.  
% The final expected log point-wise predictive density (elpd) is the sum over all observations (elpd $=\sum{\log p(y_i|y_{-i})}$).

@@

\paragraph{Model 2} $t_{p_{2}g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the early failure mode:
$$ \pi \sim \op{Logit-normal}\langle 0.007,\, 0.26 \rangle,$$
$$\sigma_1 \sim \op{Lognormal}\langle 0.14,\, 7.1\rangle,$$ 
$$t_{0.5} \sim \op{Lognormal} \langle 22,\, 5.5 \times 10^{4} \rangle$$

@@

\begin{enumerate}
\item Sample $r$ uniformly from $\{1,\ldots,S\}$.\\ For $i=1,\ldots,n_g$, do:
\item Simulate $\tilde{t}_{rep,g,i}^{(s)} \sim \op{GLFP}(\pi_{g}^{(r)},t_{p1}^{(r)},\sigma_1^{(r)},t_{p_{2}g}^{(r)}, \sigma_{2g}^{(r)})$
\item Repeat step 2 until $\tilde{t}_{rep,g,i} > t_{g,i}^L$
\item Set $d_{g,i}^{(s)}= \begin{cases} \max \{t_{g,i}: i \ge 1\}, & \mbox{ if }c_{g,i}=0\\
t_{g,i}, & \mbox{ if }c_{g, i}=1. \end{cases}$\\
Lastly, set $(t_{rep,g,i}^{(s)},c_{rep,g,i}^{(s)}) = \begin{cases}
  (d_{g,i}^{(s)},1), & \mbox{ if } \tilde{t}_{rep,g,i}^{(s)}>d_{g,i}^{(s)}\\
  (\tilde{t}_{rep,g,i}^{(s)},0), & \mbox{ if }\tilde{t}_{rep,g,i}^{(s)} \le d_{g,i}^{(s)}
\end{cases}$
\end{enumerate}
Finally, plot K-M estimates based on the $R$ replicated data sets along with the one based on the observed data on the same axes. Discrepancies between the replicated and actual data can be used to diagnose inadequacies of the model.

@@

The mean time-to-failure (MTTF) for drive-model $d$ can be estimated by
$$\frac{1}{S} \sum_{s=1}^{S} t_{g,new}^{(s)}$$
where $t_{g,new}^{(s)}$ is a draw from the posterior predictive distribution for drive-model $g$.

@@

Let $L(t) = e^{-0.4t}$ represent the value at the time of failure
relative to a new unit. We can rank the drive-models by 
$$\op{E}(L(t_{g,new})|t))\approx \frac{1}{S} \sum_{s=1}^{S} L(t_{g,new}^{(s)})$$
